---
title: "StrawberryDataCleaning"
output: pdf_document
---
1 .Reading data and ditch the counties.
```{r}
library(dplyr)
library(kableExtra)
library(tidyverse)

strawberry <- read.csv("strawberries25_v3.csv")
head(strawberry)
colnames(strawberry)
strawberry <- strawberry |> 
  filter(`Geo.Level`== "NATIONAL" | `Geo.Level`== "STATE")

```
2. Here we will drop the column with most single value, since the columns are either all missing data, or cannnot be analyzed correctly.
```{r}
drop_one_value_col <- function(df){ 
  df_id <- ensym(df)
  msg = paste("Looking for single value columns in data frame: ", as.character(df_id))
  print(msg)
  drop <- NULL
  val <- NULL

  for(i in 1:ncol(df)){   
    if(length(unique(df[[i]])) == 1){
      drop <- c(drop, i)
      val <- c(val, df[1, i])
    }
  }

  if(is.null(drop)){
    print("No columns dropped")
    return(df)
  } else {
    print("Columns dropped:")
    print(unlist(val))
    df <- df[, -drop, drop = FALSE]
    return(df)
  }
}

strawberry <- strawberry |> drop_one_value_col()

head(strawberry)
```

3.We will separate the data into census and survey, since I find the we can separate  different ways may cause some overall effect on the variable of data.item and domain-categroy.

```{r}
straw_cen <- strawberry |> filter(Program=="CENSUS")
straw_sur <- strawberry |> filter(Program=="SURVEY")

head(straw_cen)
head(straw_sur)

unique_values1 <- unique(straw_sur$Data.Item)
unique_values2 <- unique(straw_cen$Data.Item)
print(unique_values1)
```
`
4.In this step, I want to separate two interaction columns which are the domain's category which contains both domain and its category. 
```{r}
straw_cen_cleaned <- straw_cen %>%
  separate(`Data.Item`, into = c("Commodity_Type", "Operation_Measure"), sep = " - ", extra = "merge", fill = "right") %>%
  separate(Commodity_Type, into = c("Commodity", "Type"), sep = ", ", extra = "merge", fill = "right") %>%
  mutate(
    Commodity = str_trim(Commodity),
    Type = ifelse(is.na(Type), "OTHER", str_trim(Type)), 
    Operation_Measure = str_trim(Operation_Measure)
  )%>%
  dplyr::select(-Commodity)

head(straw_cen_cleaned)

unique_values3 <- unique(straw_cen_cleaned$Type)
unique_values4 <- unique(straw_cen_cleaned$Operation_Measure)
print(unique_values3)
print(unique_values4)
```
5. In this step we will focus on the survey data set. Since it is more complex, we will separate them in more columns.
```{r}
straw_sur_cleaned <- straw_sur %>%
  separate(`Data.Item`, into = c("Commodity_Market", "Details"), sep = " - ", extra = "merge", fill = "right") %>%
  separate(Commodity_Market, into = c("Commodity", "Market_Type"), sep = ", ", extra = "merge", fill = "right") %>%
  separate(Details, into = c("Measure_Operation", "Unit_of_Measure"), sep = ", MEASURED IN ", extra = "merge", fill = "right") %>%
  mutate(
    Commodity = str_trim(Commodity),
    Market_Type = ifelse(is.na(Market_Type), "OTHER", str_trim(Market_Type)),
    Measure_Operation = str_trim(Measure_Operation),
    Unit_of_Measure = str_trim(Unit_of_Measure)
  )%>%
  dplyr::select(-Commodity)

head(straw_sur_cleaned)

unique_values5 <- unique(straw_sur_cleaned$Market_Type)
unique_values6 <- unique(straw_sur_cleaned$Measure_Operation)
print(unique_values5)
print(unique_values6)
```
6: In this step I want to o the same thing on Domain.Category that separates it into two columns. I found in the survey graph, the this column has two cases, the not specified annd chemical details as mentioned in the assignment instructions.
```{r}
straw_sur_cleaned1 <- straw_sur_cleaned %>%
  separate(`Domain.Category`, into = c("Chemical_Use", "Chemical_Details"), sep = ": ", extra = "merge", fill = "right") %>%
  mutate(
    Chemical_Use = str_trim(str_replace(Chemical_Use, "CHEMICAL, ", "")),
    Chemical_Details = ifelse(Chemical_Use == "NOT SPECIFIED", "NOT SPECIFIED", Chemical_Details)
  ) %>%
  separate(Chemical_Details, into = c("Chemical_Name", "Chemical_Code"), sep = " = ", extra = "merge", fill = "right") %>%
  mutate(
    Chemical_Name = str_trim(str_replace_all(Chemical_Name, "[()]", "")), 
    Chemical_Code = str_trim(str_replace(Chemical_Code, "[)]$", ""))
  )
straw_sur_cleaned1 <- straw_sur_cleaned1 %>%
  mutate(
    Chemical_Code = as.numeric(Chemical_Code)
  )

head(straw_sur_cleaned1)
```
7. Then we will focus on the census data just like last step.
```{r}
unique_values8 <- unique(straw_cen_cleaned$Domain.Category)
print(unique_values8)
straw_cen_cleaned1 <- straw_cen_cleaned %>%
  separate(`Domain.Category`, into = c("Category_Type", "Details"), sep = ": ", extra = "merge", fill = "right") %>%
  mutate(
    Category_Type = str_trim(Category_Type),
    Details = ifelse(Category_Type == "NOT SPECIFIED", "NOT SPECIFIED", str_trim(str_replace_all(Details, "[()]", ""))) # Remove parentheses
  )

head(straw_cen_cleaned1)
```

8: In this step we will use deal with the N.A data in both data set
First: we change the NAs in State.ANSI to -1, since when we want all the variables to be numeric, which is eaiser when comparision.
```{r}
straw_sur_cleaned1 <- straw_sur_cleaned1 %>%
  mutate(
    `State.ANSI` = ifelse(is.na(`State.ANSI`), -1, `State.ANSI`)
  )
straw_cen_cleaned1 <- straw_cen_cleaned1 %>%
  mutate(
    `State.ANSI` = ifelse(is.na(`State.ANSI`), -1, `State.ANSI`)
  )
head(straw_sur_cleaned1)
head(straw_cen_cleaned1)
```
9:The second step for this is that I want to fill the missing data in the value. I will create a linear regression with Year, Category_Type, Details, State.ANSI of the census data.
```{r}
straw_cen_cleaned1 <- straw_cen_cleaned1 %>%
  mutate(Value = as.numeric(Value)) %>%
  drop_na(Value, Year, Category_Type, Details, State.ANSI)

straw_cen_cleaned1 <- straw_cen_cleaned1 %>%
  mutate(
    Category_Type = as.factor(Category_Type),
    Details = as.factor(Details)
  )

modelcen <- lm(Value ~ Year + Category_Type + Details + State.ANSI, data = straw_cen_cleaned1)

summary(modelcen)

numeric_data <- straw_cen_cleaned1 %>%
  filter(!is.na(Value) & grepl("^[0-9.]+$", Value)) %>%
  mutate(Value = as.numeric(Value))

non_numeric_data <- straw_cen_cleaned1 %>%
  filter(is.na(Value) | !grepl("^[0-9.]+$", Value))

predicted_values <- predict(modelcen, newdata = non_numeric_data)

non_numeric_data <- non_numeric_data %>%
  mutate(Value = predicted_values)

straw_cen_cleaned2 <- bind_rows(numeric_data, non_numeric_data)
head(straw_cen_cleaned2)

```
10. We will do the same thing to survey data. We found a really nice model with a R^2 over 0.58. Contrasting to the model for census data, it is more than 10 times better .
```{r}
straw_sur_cleaned1 <- straw_sur_cleaned1 %>%
  mutate(Value = as.numeric(Value)) %>%
  drop_na(Value, Year, Market_Type, Unit_of_Measure, State.ANSI)


straw_sur_cleaned1 <- straw_sur_cleaned1 %>%
  mutate(
    Market_Type = as.factor(Market_Type),
    Unit_of_Measure = as.factor(Unit_of_Measure)
  )

modelsur <- lm(Value ~ Year + Unit_of_Measure + Market_Type + State.ANSI, data = straw_sur_cleaned1)
summary(modelsur)

numeric_data <- straw_sur_cleaned1 %>%
  filter(!is.na(Value) & grepl("^[0-9.]+$", Value)) %>%
  mutate(Value = as.numeric(Value))

non_numeric_data <- straw_sur_cleaned1 %>%
  filter(is.na(Value) | !grepl("^[0-9.]+$", Value))

predicted_values <- predict(modelsur, newdata = non_numeric_data)

non_numeric_data <- non_numeric_data %>%
  mutate(Value = predicted_values)

straw_sur_cleaned2 <- bind_rows(numeric_data, non_numeric_data)
head(straw_sur_cleaned2)
```
10.We will output those two dataset.
```{r}
write.csv(straw_sur_cleaned2, "straw_sur_cleaned2.csv", row.names = FALSE)
write.csv(straw_cen_cleaned2, "straw_cen_cleaned2.csv", row.names = FALSE)
```

 